P4IRS: An Intermediate Representation and Compiler for Parallel and Performance-Portable Particle Simulations

### 2024-07-04 sage_latex_template/manuscript.tex
Adjust bibliography
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

\lstset{
    showstringspaces=false,
    extendedchars=true,
    morekeywords={self},
    emph={ParticleFor,ParticleInteraction,Lowerable,Assign},
    basicstyle=\footnotesize\ttfamily,
    commentstyle=\slshape\color{deepgreen},
    stringstyle=\ttfamily\color{deepred},
    keywordstyle=\ttfamily\color{deepblue},
    emphstyle=\ttfamily\color{violet},
    breaklines=true,
    breakatwhitespace=true,
    columns=flexible,
    numbers=left,
    numberstyle=\tiny,
    basewidth=.5em,
    xleftmargin=.5cm,
    captionpos=b,
    frame=lines
}

@@

P4IRS can generate several kernel flavours of the same potential such as Linked Cells \cite{linkedcells} and Verlet Lists \cite{verletlists}, where data layouts for properties and specific arrays (such as the neighbor-lists) can be switched to the case with best performance.
A runtime interface in C++ is also available and can be used to integrate P4IRS with other simulation tools, which is particularly important for integrating P4IRS simulations with our multi-physics framework waLBerLa \cite{walberla1,walberla2}.
%as an example we show how P4IRS can use the load-balancing communication algorithm from waLBerLa to compute simulations with a more heterogeneous distribution of particles.

@@

The goals of this paper are summarized as follows:
\begin{itemize}
    \item We present our code generator tool P4IRS, describing how it works and how it can be used to generate efficient particle simulation codes.
    \item We discuss the advantages of using P4IRS to achieve performance-portability, also comparing it with other previous approaches such as tinyMD \cite{tinymd} and MESA-PD \cite{mesapd1,mesapd2}.
    \item We describe P4IRS interfaces to allow its generated code to be integrated with other simulation tools, specifically in the context of domain-partitioning and communication.
    \item We provide performance results for Intel Ice Lake and AMD Milan CPUs, as well as for NVIDIA A40 and A100 accelerators to demonstrate the state-of-the-art performance of the applications generated by P4IRS in comparison to MD-Bench and MESA-PD.
    \item We display weak-scaling results in the Fritz and JUWELS Booster clusters to demonstrate the weak-scaling capabilities of P4IRS in MD and DEM simulations on modern supercomputers.
\end{itemize}

@@

In MD, there exist many packages with distinct strategies and simulation fields.
LAMMPS \cite{lammps1,lammps2} is a molecular package for the material-modeling field which uses the standard Verlet Lists algorithm to optimize the non-bonded short-range force calculations.
It offers many inter-atomic potentials, each one with different kernel implementations for a specific target (like CPU, GPU or Kokkos \cite{kokkos}).
GROMACS \cite{gromacs1,gromacs2} is another package for MD with focus on bio-sciences simulation which uses its own Cluster Pair algorithm to optimize the short-range forces computation.
It does not support as many potentials as LAMMPS, but its algorithm tends to be significantly faster than the Verlet Lists since it employs SIMD parallelism more efficiently.
Furthermore, the optimal GROMACS kernel implementations rely on SIMD intrinsics to attain higher performance, which makes them harder to be maintained and ported to other architectures.
Also, the GPU algorithm in GROMACS differs from the CPU, since it relies on a strategy named ``super-clustering'' to reduce the amount of redundant computations caused by adjusting the cluster sizes to fit the SIMD width in the GPU.
Other MD packages such as HOOMD-blue \cite{hoomdblue}, NAMD \cite{namd}, AMBER \cite{amber1,amber2} and CHARMM \cite{charmm} also contain dedicated and hand-tuned implementations in general purpose languages for each supported backend.
Each version is tested, maintained, debugged and optimized independently.

Packages that support DEM simulations are in general more flexible (since simulation with more particle properties such as rotations and shapes are needed), however they usually do not have the same optimization techniques as the MD simulation packages.
The reason is most likely that it is not so feasible to support flexibility with sophisticated optimization algorithms as the complexity of the code grows considerably.
Examples of such packages are the LIGGGHTS DEM \cite{liggghts} (based on LAMMPS), MercuryDPM \cite{mercurydpm} and YADE \cite{yade}.
To mitigate such limitations, a few code generation approaches were developed for both MD and DEM.

@@

OpenMM \cite{openmm} is a \ac{MD} package that makes use of Python as a front-end for simulation protocols and file I/O, it contains a C++ API for forces and integrators as well as optimized kernel implementations for CUDA, C++ and OpenCL.
Integrators and forces can be implemented by using mathematical descriptions in OpenMM, but its use cases are restricted to \ac{MD} simulations and several features required for a DEM simulation or other types of particles simulation (such as contact history and different shapes) are not available.

MESA-PD \cite{mesapd3} is a general particle dynamics framework that is built-in the waLBerla framework for the simulation of particles and fluids.
Its design principle of separation of code and data allows to introduce arbitrary interaction models with ease, which makes it also useful for \ac{MD}.
By using its code generation approach using Jinja templates it can be adapted very effectively to any simulation scenario.
As successor of the ``physics engine'' (pe) rigid particle dynamics framework it inherits its scalability and advanced load balancing functionalities.
One of the limitations of MESA-PD is that it does not support GPU code generation, and since its code generation approach is based on Jinja templates, the effort for implementing new functionalities may not be lower compared to other simulation packages without code generation techniques (for instance, most of its force kernels are not generated but rather implemented in C++).

In previous work we developed tinyMD, a proxy-app (also based on miniMD) created to evaluate the portability of MD and DEM applications with the AnyDSL \cite{anydsl1,anydsl2} framework.
TinyMD uses higher-order functions to abstract device iteration loops, data layouts and communication strategies, providing a domain-specific library to implement pair-wise interaction kernels that execute efficiently on multi-CPU and multi-GPU targets.
One of tinyMD's limitation is that its flexibility and optimization settings are restricted to what can be achieve through higher-order functions and the AnyDSL compilation framework (the generated LLVM IR restricts the backend compilation to the Clang compiler), overcoming such limitation is one of the reasons we develop P4IRS.

@@

Many performance-oriented researches exist for MD systems, most of them using MD packages as a black-box and adjusting different simulation parameters to analyse their implications in the overall performance.
In some cases, however, in-depth performance studies were made to gain a broader understanding on efficient ways to achieve SIMD vectorization with SSE/AVX instructions \cite{mdsimd}.
Most in-depth studies are performed using mini-apps or proxy-apps, one example is miniMD, a proxy-app that provides a simple copper face-cubic-centered lattice simulation where atom forces are computed with kernels extracted from LAMMPS for either the \ac{LJ} or \ac{EAM} potentials.

The Exascale Computing Project (ECP) Co-design Center for Particle Applications (CoPA) \cite{ecpcopa} focus on co-designing particle applications to prepare them for exascale computing.
They provide several proxy-apps for particle simulations such as CabanaMD and ExaMiniMD, which are designed in a modular fashion to account for different optimization strategies and particle data layouts, and also are built on top of Kokkos to be able to execute on many-core CPU and GPU targets.

Our performance-focused prototyping harness MD-Bench \cite{mdbench1,mdbench2} is a toolbox developed for performance-engineering \ac{MD} non-bonded short-range force calculation algorithms, with focus on in-core execution on CPUs and GPUs.
It currently supports the Verlet Lists and Cluster Pair algorithms from LAMMPS and GROMACS, respectively.
One of our goals is to use the knowledge obtained with MD-Bench and integrate it into P4IRS to achieve state-of-the-art performance and flexibility altogether in particle simulations.

@@

Where $k^{n}$ is the normal spring stiffness, $k^{t}$ the tangential spring stiffness, $\xi^{n}$ the normal velocity damping, $\mu$ the friction, $\delta_{ij}$ the penetration depth and $dt$ the time-step size.
The update rule in \autoref{eq:update_rule} and modifications in \autoref{eq:linear_spring_dashpot_tangential} were properly included to fix frictional contact accordingly \cite{dem1,dem2}.
The tangential spring displacement $\hat{\delta}^{t}_{ij}$ is initialized as a zero vector in the first contact between the particles $i$ and $j$, and the update rule is carried out every time-step the contact persists.
From the algorithmic point of view, this means that a contact history data structure is required to keep track of the last computed value for every pair of particles that are in contact.

@@

The icc (ICC) 2021.10.0 20230609 compiler was used for CPU experiments (except in the JUWELS cluster), with the \emph{-Ofast -xCORE-AVX512 -qopt-zmm-usage=high} flags for Fritz / Ice Lake (AVX512) and \emph{-Ofast -march=core-avx2} flags for Milan (AVX2).
For CUDA, we use the NVCC compiler V12.1.r12.1 for the single GPU experiments, and the same flags used for Milan (AVX2) since it is the host processor in the systems with the A40/A100 GPUs.
In the JUWELS Booster, we used the Intel(R) oneAPI DPC++/C++ Compiler 2023.2.0 (2023.2.0.20230721) together with NVCC V12.2.91, and the same flags as used for Milan since both CPUs have AVX2 with FMA as the largest SIMD vector extension available.
We make use of the Likwid toolkit V5.3 \cite{likwid} to pin tasks to specific CPU cores and measure HPM counters in the target CPU (Ice Lake) for the single core experiments, where P4IRS is able to properly generate the instrumentation markers and MD-Bench kernels are already instrumented.

@@

%\bibliographystyle{elsarticle-num-names} 
\bibliographystyle{plain}
\bibliography{literature}

%\begin{thebibliography}{99}
%\bibitem[Kopka and Daly(2003)]{R1}
%Kopka~H and Daly~PW (2003) \textit{A Guide to \LaTeX}, 4th~edn.
%Addison-Wesley.

%\bibitem[Lamport(1994)]{R2}
%Lamport~L (1994) \textit{\LaTeX: a Document Preparation System},
%2nd~edn. Addison-Wesley.

@@

%\end{thebibliography}

>>>>

\lstset{
    showstringspaces=false,
    extendedchars=true,
    morekeywords={self},
    emph={ParticleFor,ParticleInteraction,Assign},
    basicstyle=\footnotesize\ttfamily,
    commentstyle=\slshape\color{deepgreen},
    stringstyle=\ttfamily\color{deepred},
    keywordstyle=\ttfamily\color{deepblue},
    emphstyle=\ttfamily\color{violet},
    breaklines=true,
    breakatwhitespace=true,
    columns=flexible,
    numbers=left,
    numberstyle=\tiny,
    basewidth=.5em,
    xleftmargin=.5cm,
    captionpos=b,
    frame=lines
}

@@

P4IRS can generate several kernel flavours of the same potential such as Linked Cells (\cite{linkedcells}) and Verlet Lists (\cite{verletlists}), where data layouts for properties and specific arrays (such as the neighbor-lists) can be switched to the case with best performance.
A runtime interface in C++ is also available and can be used to integrate P4IRS with other simulation tools, which is particularly important for integrating P4IRS simulations with our multi-physics framework waLBerLa (\cite{walberla1,walberla2}).
%as an example we show how P4IRS can use the load-balancing communication algorithm from waLBerLa to compute simulations with a more heterogeneous distribution of particles.

@@

The goals of this paper are summarized as follows:
\begin{itemize}
    \item We present our code generator tool P4IRS, describing how it works and how it can be used to generate efficient particle simulation codes.
    \item We discuss the advantages of using P4IRS to achieve performance-portability, also comparing it with other previous approaches such as tinyMD (\cite{tinymd}) and MESA-PD (\cite{mesapd1,mesapd2}).
    \item We describe P4IRS interfaces to allow its generated code to be integrated with other simulation tools, specifically in the context of domain-partitioning and communication.
    \item We provide performance results for Intel Ice Lake and AMD Milan CPUs, as well as for NVIDIA A40 and A100 accelerators to demonstrate the state-of-the-art performance of the applications generated by P4IRS in comparison to MD-Bench and MESA-PD.
    \item We display weak-scaling results in the Fritz and JUWELS Booster clusters to demonstrate the weak-scaling capabilities of P4IRS in MD and DEM simulations on modern supercomputers.
\end{itemize}

@@

In MD, there exist many packages with distinct strategies and simulation fields.
LAMMPS (\cite{lammps1,lammps2}) is a molecular package for the material-modeling field which uses the standard Verlet Lists algorithm to optimize the non-bonded short-range force calculations.
It offers many inter-atomic potentials, each one with different kernel implementations for a specific target such as CPU, GPU or Kokkos (\cite{kokkos}).
GROMACS (\cite{gromacs1,gromacs2}) is another package for MD with focus on bio-sciences simulation which uses its own Cluster Pair algorithm to optimize the short-range forces computation.
It does not support as many potentials as LAMMPS, but its algorithm tends to be significantly faster than the Verlet Lists since it employs SIMD parallelism more efficiently.
Furthermore, the optimal GROMACS kernel implementations rely on SIMD intrinsics to attain higher performance, which makes them harder to be maintained and ported to other architectures.
Also, the GPU algorithm in GROMACS differs from the CPU, since it relies on a strategy named ``super-clustering'' to reduce the amount of redundant computations caused by adjusting the cluster sizes to fit the SIMD width in the GPU.
Other MD packages such as HOOMD-blue (\cite{hoomdblue}), NAMD (\cite{namd}), AMBER (\cite{amber1}) and CHARMM (\cite{charmm}) also contain dedicated and hand-tuned implementations in general purpose languages for each supported backend.
Each version is tested, maintained, debugged and optimized independently.

Packages that support DEM simulations are in general more flexible (since simulation with more particle properties such as rotations and shapes are needed), however they usually do not have the same optimization techniques as the MD simulation packages.
The reason is most likely that it is not so feasible to support flexibility with sophisticated optimization algorithms as the complexity of the code grows considerably.
Examples of such packages are the LIGGGHTS DEM (\cite{liggghts}) (based on LAMMPS), MercuryDPM (\cite{mercurydpm}) and YADE (\cite{yade}).
To mitigate such limitations, a few code generation approaches were developed for both MD and DEM.

@@

OpenMM (\cite{openmm}) is a \ac{MD} package that makes use of Python as a front-end for simulation protocols and file I/O, it contains a C++ API for forces and integrators as well as optimized kernel implementations for CUDA, C++ and OpenCL.
Integrators and forces can be implemented by using mathematical descriptions in OpenMM, but its use cases are restricted to \ac{MD} simulations and several features required for a DEM simulation or other types of particles simulation (such as contact history and different shapes) are not available.

MESA-PD (\cite{mesapd3}) is a general particle dynamics framework that is built-in the waLBerla framework for the simulation of particles and fluids.
Its design principle of separation of code and data allows to introduce arbitrary interaction models with ease, which makes it also useful for \ac{MD}.
By using its code generation approach using Jinja templates it can be adapted very effectively to any simulation scenario.
As successor of the ``physics engine'' (pe) rigid particle dynamics framework it inherits its scalability and advanced load balancing functionalities.
One of the limitations of MESA-PD is that it does not support GPU code generation, and since its code generation approach is based on Jinja templates, the effort for implementing new functionalities may not be lower compared to other simulation packages without code generation techniques (for instance, most of its force kernels are not generated but rather implemented in C++).

In previous work we developed tinyMD, a proxy-app (also based on miniMD) created to evaluate the portability of MD and DEM applications with the AnyDSL (\cite{anydsl1,anydsl2}) framework.
TinyMD uses higher-order functions to abstract device iteration loops, data layouts and communication strategies, providing a domain-specific library to implement pair-wise interaction kernels that execute efficiently on multi-CPU and multi-GPU targets.
One of tinyMD's limitation is that its flexibility and optimization settings are restricted to what can be achieve through higher-order functions and the AnyDSL compilation framework (the generated LLVM IR restricts the backend compilation to the Clang compiler), overcoming such limitation is one of the reasons we develop P4IRS.

@@

Many performance-oriented researches exist for MD systems, most of them using MD packages as a black-box and adjusting different simulation parameters to analyse their implications in the overall performance.
In some cases, however, in-depth performance studies were made to gain a broader understanding on efficient ways to achieve SIMD vectorization with SSE/AVX instructions (\cite{mdsimd}).
Most in-depth studies are performed using mini-apps or proxy-apps, one example is miniMD, a proxy-app that provides a simple copper face-cubic-centered lattice simulation where atom forces are computed with kernels extracted from LAMMPS for either the \ac{LJ} or \ac{EAM} potentials.

The Exascale Computing Project (ECP) Co-design Center for Particle Applications (CoPA) (\cite{ecpcopa}) focus on co-designing particle applications to prepare them for exascale computing.
They provide several proxy-apps for particle simulations such as CabanaMD and ExaMiniMD, which are designed in a modular fashion to account for different optimization strategies and particle data layouts, and also are built on top of Kokkos to be able to execute on many-core CPU and GPU targets.

Our performance-focused prototyping harness MD-Bench (\cite{mdbench1,mdbench2}) is a toolbox developed for performance-engineering \ac{MD} non-bonded short-range force calculation algorithms, with focus on in-core execution on CPUs and GPUs.
It currently supports the Verlet Lists and Cluster Pair algorithms from LAMMPS and GROMACS, respectively.
One of our goals is to use the knowledge obtained with MD-Bench and integrate it into P4IRS to achieve state-of-the-art performance and flexibility altogether in particle simulations.

@@

Where $k^{n}$ is the normal spring stiffness, $k^{t}$ the tangential spring stiffness, $\xi^{n}$ the normal velocity damping, $\mu$ the friction, $\delta_{ij}$ the penetration depth and $dt$ the time-step size.
The update rule in \autoref{eq:update_rule} and modifications in \autoref{eq:linear_spring_dashpot_tangential} were properly included to fix frictional contact accordingly (\cite{dem1,dem2}).
The tangential spring displacement $\hat{\delta}^{t}_{ij}$ is initialized as a zero vector in the first contact between the particles $i$ and $j$, and the update rule is carried out every time-step the contact persists.
From the algorithmic point of view, this means that a contact history data structure is required to keep track of the last computed value for every pair of particles that are in contact.

@@

The icc (ICC) 2021.10.0 20230609 compiler was used for CPU experiments (except in the JUWELS cluster), with the \emph{-Ofast -xCORE-AVX512 -qopt-zmm-usage=high} flags for Fritz / Ice Lake (AVX512) and \emph{-Ofast -march=core-avx2} flags for Milan (AVX2).
For CUDA, we use the NVCC compiler V12.1.r12.1 for the single GPU experiments, and the same flags used for Milan (AVX2) since it is the host processor in the systems with the A40/A100 GPUs.
In the JUWELS Booster, we used the Intel(R) oneAPI DPC++/C++ Compiler 2023.2.0 (2023.2.0.20230721) together with NVCC V12.2.91, and the same flags as used for Milan since both CPUs have AVX2 with FMA as the largest SIMD vector extension available.
We make use of the Likwid toolkit V5.3 (\cite{likwid}) to pin tasks to specific CPU cores and measure HPM counters in the target CPU (Ice Lake) for the single core experiments, where P4IRS is able to properly generate the instrumentation markers and MD-Bench kernels are already instrumented.

@@

\bibliographystyle{plain}

\begin{thebibliography}{99}

@@

\bibitem[Anderson~J~A et al.(2020)]{hoomdblue}
Joshua~A. Anderson, Jens Glaser, and Sharon~C. Glotzer.
\newblock Hoomd-blue: A python package for high-performance molecular dynamics
  and hard particle monte carlo simulations.
\newblock {\em Computational Materials Science}, 173:109363, 2020.

\bibitem[Brendel~L and Dippel~S(1998)]{dem1}
L.~Brendel and S.~Dippel.
\newblock {\em Lasting Contacts in Molecular Dynamics Simulations}, pages
  313--318.
\newblock Springer Netherlands, Dordrecht, 1998.

\bibitem[Brooks~B~R et al.(2009)]{charmm}
B~R Brooks, C~L Brooks, A~D Mackerell, L~Nilsson, R~J Petrella, B~Roux, Y~Won,
  G~Archontis, C~Bartels, S~Boresch, A~Caflisch, L~Caves, Q~Cui, A~R Dinner,
  M~Feig, S~Fischer, J~Gao, M~Hodoscek, W~Im, K~Kuczera, T~Lazaridis, J~Ma,
  V~Ovchinnikov, E~Paci, R~W Pastor, C~B Post, J~Z Pu, M~Schaefer, B~Tidor, R~M
  Venable, H~L Woodcock, X~Wu, W~Yang, D~M York, and M~Karplus.
\newblock Charmm: the biomolecular simulation program.
\newblock {\em Journal of Computational Chemistry}, 30(10):1545--1614, 2009.

\bibitem[Brown~W et al.(2012)]{lammps2}
W.~Brown, Axel Kohlmeyer, Steven Plimpton, and Arnold Tharrington.
\newblock Implementing molecular dynamics on hybrid high performance computers
  - particle-particle particle-mesh.
\newblock {\em Computer Physics Communications}, 183:449--459, 03 2012.

\bibitem[{Carter Edwards}~H et al.(2014)]{kokkos}
H.~{Carter Edwards}, Christian~R. Trott, and Daniel Sunderland.
\newblock Kokkos: Enabling manycore performance portability through polymorphic
  memory access patterns.
\newblock {\em Journal of Parallel and Distributed Computing},
  74(12):3202--3216, 2014.
\newblock Domain-Specific Languages and High-Level Frameworks for
  High-Performance Computing.

\bibitem[Case~D~A. et al.(2023)]{amber1}
David~A. Case, Hasan~Metin Aktulga, Kellon Belfon, David~S. Cerutti, G.~Andrés
  Cisneros, Vinícius Wilian~D. Cruzeiro, Negin Forouzesh, Timothy~J. Giese,
  Andreas~W. Götz, Holger Gohlke, Saeed Izadi, Koushik Kasavajhala, Mehmet~C.
  Kaymak, Edward King, Tom Kurtzman, Tai-Sung Lee, Pengfei Li, Jian Liu, Tyler
  Luchko, Ray Luo, Madushanka Manathunga, Matias~R. Machado, Hai~Minh Nguyen,
  Kurt~A. O’Hearn, Alexey~V. Onufriev, Feng Pan, Sergio Pantano, Ruxi Qi, Ali
  Rahnamoun, Ali Risheh, Stephan Schott-Verdugo, Akhil Shajan, Jason Swails,
  Junmei Wang, Haixin Wei, Xiongwu Wu, Yongxian Wu, Shi Zhang, Shiji Zhao,
  Qiang Zhu, Thomas E.~III Cheatham, Daniel~R. Roe, Adrian Roitberg, Carlos
  Simmerling, Darrin~M. York, Maria~C. Nagan, and Kenneth M.~Jr. Merz.
\newblock Ambertools.
\newblock {\em Journal of Chemical Information and Modeling},
  63(20):6183--6191, 2023.
\newblock PMID: 37805934.

\bibitem[Eastman~P et al.(2017)]{openmm}
Peter Eastman, Jason Swails, John~D. Chodera, Robert~T. McGibbon, Yutong Zhao,
  Kyle~A. Beauchamp, Lee-Ping Wang, Andrew~C. Simmonett, Matthew~P. Harrigan,
  Chaya~D. Stern, Rafal~P. Wiewiora, Bernard~R. Brooks, and Vijay~S. Pande.
\newblock Openmm 7: Rapid development of high performance algorithms for
  molecular dynamics.
\newblock {\em PLOS Computational Biology}, 13(7):1--17, 07 2017.

\bibitem[Eibl~S and Rüde~U(2019)]{mesapd3}
Sebastian Eibl and Ulrich R{\"{u}}de.
\newblock A modular and extensible software architecture for particle dynamics.
\newblock {\em Proceedings of the 8\textsuperscript{th} International
  Conference on Discrete Element Methods (DEM8)}, 2019.

\bibitem[Eibl~S and Rüde~U(2018)]{mesapd1}
Sebastian Eibl and Ulrich Rüde.
\newblock A local parallel communication algorithm for polydisperse rigid body
  dynamics.
\newblock {\em Parallel Computing}, 80:36--48, 2018.

\bibitem[Eibl~S and Rüde~U(2019)]{mesapd2}
Sebastian Eibl and Ulrich Rüde.
\newblock A systematic comparison of runtime load balancing algorithms for
  massively parallel rigid particle dynamics.
\newblock {\em Computer Physics Communications}, 244:76--85, 2019.

\bibitem[Godenschwager~C et al.(2013)]{walberla1}
Christian Godenschwager, Florian Schornbaum, Martin Bauer, Harald K\"{o}stler,
  and Ulrich R\"{u}de.
\newblock A framework for hybrid parallel flow simulations with a trillion
  cells in complex geometries.
\newblock In {\em Proceedings of the International Conference on High
  Performance Computing, Networking, Storage and Analysis}, SC '13, New York,
  NY, USA, 2013. Association for Computing Machinery.

\bibitem[Khakhar~D~V et al.(1997)]{dem2}
D.~V. Khakhar, J.~J. McCarthy, and J.~M. Ottino.
\newblock {Radial segregation of granular mixtures in rotating cylinders}.
\newblock {\em Physics of Fluids}, 9(12):3600--3614, 12 1997.

\bibitem[Kloss~C et al.(2012)]{liggghts}
Christoph Kloss, Christoph Goniva, Alice König, Stefan Amberger, and Stefan
  Pirker.
\newblock Models, algorithms and validation for opensource dem and cfd-dem.
\newblock {\em Progress in Computational Fluid Dynamics}, 12:140 -- 152, 06
  2012.

\bibitem[Lei{\ss}a~R et al.(2018)]{anydsl1}
Roland Lei{\ss}a, Klaas Boesche, Sebastian Hack, Ars{\`{e}}ne
  P{\'{e}}rard{-}Gayot, Richard Membarth, Philipp Slusallek, Andr{\'{e}}
  M{\"{u}}ller, and Bertil Schmidt.
\newblock {AnyDSL}: A partial evaluation framework for programming
  high-performance libraries.
\newblock {\em {PACMPL}}, 2({OOPSLA}):119:1--119:30, 2018.

\bibitem[Lei{\ss}a~R et al.(2015)]{anydsl2}
Roland Lei{\ss}a, Marcel K{\"{o}}ster, and Sebastian Hack.
\newblock A graph-based higher-order intermediate representation.
\newblock In {\em Proceedings of the 13th Annual {IEEE/ACM} International
  Symposium on Code Generation and Optimization, {CGO} 2015, San Francisco, CA,
  USA, February 07 - 11, 2015}, pages 202--212, 2015.

\bibitem[Machado~R~R~L et al.(2021)]{tinymd}
Rafael Ravedutti~L. Machado, Jonas Schmitt, Sebastian Eibl, Jan Eitzinger,
  Roland Leißa, Sebastian Hack, Arsène Pérard-Gayot, Richard Membarth, and
  Harald Köstler.
\newblock tinymd: Mapping molecular dynamics simulations to heterogeneous
  hardware using partial evaluation.
\newblock {\em Journal of Computational Science}, 54:101425, 2021.

\bibitem[Mniszewski~S~M et al.(2021)]{ecpcopa}
Susan~M Mniszewski, James Belak, Jean-Luc Fattebert, Christian~FA Negre,
  Stuart~R Slattery, Adetokunbo~A Adedoyin, Robert~F Bird, Choongseok Chang,
  Guangye Chen, Stéphane Ethier, Shane Fogerty, Salman Habib, Christoph
  Junghans, Damien Lebrun-Grandié, Jamaludin Mohd-Yusof, Stan~G Moore, Daniel
  Osei-Kuffuor, Steven~J Plimpton, Adrian Pope, Samuel~Temple Reeve, Lee
  Ricketson, Aaron Scheinberg, Amil~Y Sharma, and Michael~E Wall.
\newblock Enabling particle applications for exascale computing platforms.
\newblock {\em The International Journal of High Performance Computing
  Applications}, 35(6):572–597, July 2021.

\bibitem[P{\'a}ll~S et al.(2015)]{gromacs2}
Szil{\'a}rd P{\'a}ll, Mark~James Abraham, Carsten Kutzner, Berk Hess, and Erik
  Lindahl.
\newblock Tackling exascale software challenges in molecular dynamics
  simulations with gromacs.
\newblock In Stefano Markidis and Erwin Laure, editors, {\em Solving Software
  Challenges for Exascale}, pages 3--27, Cham, 2015. Springer International
  Publishing.

\bibitem[Pennycook~J et al.(2013)]{mdsimd}
John Pennycook, Chris Hughes, M.~Smelyanskiy, and Stephen Jarvis.
\newblock Exploring simd for molecular dynamics, using intel® xeon®
  processors and intel® xeon phi coprocessors.
\newblock In {\em Proceedings - IEEE 27th International Parallel and
  Distributed Processing Symposium, IPDPS 2013}, pages 1085--1097, 05 2013.

\bibitem[Phillips~J~C et al.(2005)]{namd}
James~C. Phillips, Rosemary Braun, Wei Wang, James Gumbart, Emad Tajkhorshid,
  Elizabeth Villa, Christophe Chipot, Robert~D. Skeel, Laxmikant Kalé, and
  Klaus Schulten.
\newblock Scalable molecular dynamics with namd.
\newblock {\em Journal of Computational Chemistry}, 26(16):1781--1802, 2005.

\bibitem[Plimpton~S(1995)]{lammps1}
Steve Plimpton.
\newblock Fast parallel algorithms for short-range molecular dynamics.
\newblock {\em Journal of Computational Physics}, 117(1):1--19, 1995.

\bibitem[Páll~S and Hess~B(2013)]{gromacs1}
Szilárd Páll and Berk Hess.
\newblock A flexible algorithm for calculating pair interactions on simd
  architectures.
\newblock {\em Computer Physics Communications}, 184(12):2641--2650, 2013.

\bibitem[Quentrec~B and Brot~C(1973)]{linkedcells}
B~Quentrec and C~Brot.
\newblock New method for searching for neighbors in molecular dynamics
  computations.
\newblock {\em Journal of Computational Physics}, 13(3):430--432, 1973.

\bibitem[Ravedutti Lucio Machado~R et al.(2023)]{mdbench1}
Rafael Ravedutti Lucio~Machado, Jan Eitzinger, Harald K{\"o}stler, and Gerhard
  Wellein.
\newblock MD-Bench: A generic proxy-app toolbox for state-of-the-art molecular
  dynamics algorithms.
\newblock In Roman Wyrzykowski, Jack Dongarra, Ewa Deelman, and Konrad
  Karczewski, editors, {\em Parallel Processing and Applied Mathematics}, pages
  321--332, Cham, 2023. Springer International Publishing.

\bibitem[Ravedutti Lucio Machado~R et al.(2023)b]{mdbench2}
Rafael {Ravedutti Lucio Machado}, Jan Eitzinger, Jan Laukemann, Georg Hager,
  Harald Köstler, and Gerhard Wellein.
\newblock MD-Bench: A performance-focused prototyping harness for
  state-of-the-art short-range molecular dynamics algorithms.
\newblock {\em Future Generation Computer Systems}, 149:25--38, 2023.

\bibitem[Schornbaum~F and R\"{u}de~U(2016)]{walberla2}
Florian Schornbaum and Ulrich R\"{u}de.
\newblock Massively parallel algorithms for the lattice boltzmann method on
  nonuniform grids.
\newblock {\em SIAM Journal on Scientific Computing}, 38(2):C96--C126, 2016.

\bibitem[Smilauer~V et al.(2021)]{yade}
Vaclav Smilauer, Vasileios Angelidakis, Emanuele Catalano, Robert Caulk, Bruno
  Chareyre, William Chèvremont, Sergei Dorofeenko, Jerome Duriez, Nolan Dyck,
  Jan Elias, Burak Er, Alexander Eulitz, Anton Gladky, Ning Guo, Christian
  Jakob, Francois Kneib, Janek Kozicki, Donia Marzougui, Raphael Maurin, Chiara
  Modenese, Gerald Pekmezi, Luc Scholtès, Luc Sibille, Jan Stransky, Thomas
  Sweijen, Klaus Thoeni, and Chao Yuan.
\newblock {\em Yade documentation}.
\newblock The Yade Project, November 2021.

\bibitem[Treibig~J et al.(2010)]{likwid}
Jan Treibig, Georg Hager, and Gerhard Wellein.
\newblock Likwid: A lightweight performance-oriented tool suite for x86
  multicore environments.
\newblock In {\em 2010 39th International Conference on Parallel Processing
  Workshops}, pages 207--216, 2010.

\bibitem[Verlet~L(1967)]{verletlists}
Loup Verlet.
\newblock Computer "experiments" on classical fluids. i. thermodynamical
  properties of lennard-jones molecules.
\newblock {\em Phys. Rev.}, 159:98--103, Jul 1967.

\bibitem[Weinhart~T et al.(2020)]{mercurydpm}
Thomas Weinhart, Luca Orefice, Mitchel Post, Marnix~P. {van Schrojenstein
  Lantman}, Irana~F.C. Denissen, Deepak~R. Tunuguntla, J.M.F. Tsang, Hongyang
  Cheng, Mohamad~Yousef Shaheen, Hao Shi, Paolo Rapino, Elena Grannonio, Nunzio
  Losacco, Joao Barbosa, Lu~Jing, Juan~E. {Alvarez Naranjo}, Sudeshna Roy,
  Wouter~K. {den Otter}, and Anthony~R. Thornton.
\newblock Fast, flexible particle simulations — an introduction to
  mercurydpm.
\newblock {\em Computer Physics Communications}, 249:107129, 2020.

\end{thebibliography}
